Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-02 doc/RMD/RDA-analysis.Rmd
add original analysis to RDA to remove collinearity

```{r resultplot, fig.width = 5, fig.height = 5}
set.seed(999)
plot_dbrda(cap11cc, arrows = 5) + 
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
# if (!interactive()){
  FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
# }
plot_dbrda(cap16cc, arrows = 10) +
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
```

This is all well and good, but what exactly does it actually mean to have only 
1.93 variance explained (for 11 loci, that is)? I was a bit stumped as well. 
Luckily, Gavin Simpson has some answers: http://stackoverflow.com/a/22537820/2752888

Basically, we have a situation where we have all of the variables only
explaining a whopping 16 - 17% of the variance and the rest goes unexplained.

@@

Out of both of the models, only Region explains the most variance, almost double
that of Host or Year and Severity contributes almost nothing.

```{r}
wmn11 <- seppop(dat11, ~Source)$wmn %>% setPop(~Region)
wmn11_sub <- popsub(wmn11, blacklist = character(0))
min_span_net <- bruvo.msn(wmn11_sub, replen = other(wmn11)$REPLEN, add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(67)
plot_poppr_msn(wmn11,
               min_span_net,
               inds = "none",
               mlg = TRUE,
               gadj = 6,
               nodelab = 100,
               nodebase = 1.15,
               palette = funky,
               cutoff = (0.5/11)*3,
               quantiles = FALSE,
               beforecut = TRUE,
               layfun = igraph::layout_with_lgl)
```

>>>>

```{r resultplot, fig.width = 5, fig.height = 5}
set.seed(999)
plot_dbrda(cap11cc, arrows = 5) + 
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
plot_dbrda(cap16cc, arrows = 10) +
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
```

This is all well and good, but what exactly does it actually mean to have only 
3.33% variance explained (for 11 loci, that is) on the first axis? I was a bit stumped as well. 
Luckily, Gavin Simpson has some answers: http://stackoverflow.com/a/22537820/2752888

Basically, we have a situation where we have all of the variables
explaining 45% of the variance and the rest goes unexplained.

@@

Out of both of the models, MCG explains the most variance, but we have an 
interesting warning pop up indicating that there is multicollinearity within our
variables. It appears that all of these are multicollinear with MCG. Since dbRDA
assumes that all the variables are independent, it would be prudent to re-run 
the analysis without the MCG variables as well

## Removing Source of Multicollinearity

```{r rda_calculation2, cache = TRUE}
# 11 loci
ENV112   <- select(ENV11, -MCG)
cap11cc2 <- choose_dbrda(dat11cc.bruvo, ENV = ENV112, CHOOSER = "ordistep")
```

```{r}
cap11cc2
cap11cc2$anova
vegan::RsquareAdj(cap11cc2)
```

```{r total_anova2, cache = TRUE}
anova(cap11cc2)
```

```{r marginal_anova2, cache = TRUE}
anova(cap11cc2, by = "margin")
```

```{r resultplot2, fig.width = 5, fig.height = 5}
set.seed(999)
plot_dbrda(cap11cc2, arrows = 5) + 
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
if (!interactive()){
  FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
}
```

What we can see here is that the amount of explained variance went down a bit to
1.92, and our model now only explains 16% of the variance.

```{r, echo = FALSE, results = "asis"}
cap11sum2 <- vegan:::summary.cca(cap11cc2, axes = 0, scaling = "sites")
knitr::kable(vartab(cap11sum2), digits = 3)
```

```{r}
vp11     <- varpart(dat11cc.bruvo, ~Severity + Year + Region + Host, data = ENV11, comm = dat11raw, add = TRUE)
plot(vp11, Xnames = c("Full Model", "No Model"))
vp11 <- varpart(dat11cc.bruvo, ~Severity, ~Year, ~Region, ~Host, data = ENV11, add = TRUE)

plot(vp11, digits = 2, Xnames = c("Year", "Region", "Host", "MCG"))
```

# Conclusions

This analysis showed us two things: Severity is not an important predictor for 
genetic diversity, but Region, Host, Year, and MCG are all predictors of genetic
diversit. Explanations for this result could be due to a few reasons: MCG, a
phenotypic variable associated with pre-zygotic barriers has SOME genetic 
component to it; the MCGs that explained the most variation were also the MCGs
that were strongly associated with one genotype (i.e. MCG 5 has 19 isolates of
MLG 25).

### 2017-08-02 doc/manuscript/manuscript.Rmd
add original analysis to RDA to remove collinearity

A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
We used this method to delineate the phenotypic, geographic, and temporal components in influencing genetic composition of the populations.
<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr]. 
<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis. -->
To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

<!-- 
 - Big Picture reminder of importance
   + This is an extension of a previous work [@otto-hanson2011variation]
   + White Mold Nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
   + It's unclear how genetic structure can influence
 - Critical Analysis of Major Findings
   + This is the largest genetic analysis of *S. sclerotiorum* to date
   + Similar to the findings of @otto-hanson2011variation, we found that
     Region has a strong influence on both population structure and
     severity.
   + North Dakota appeared
 - Additional Findings and how they fit into the literature
 - Limitations
 - Future Directions
 - Overall Conclusion and Impact
-->

@@

Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (Aggressiveness, Region, Source,   Year, and Host) [@legendre1999distance].
From the initial results, it appears that the most important factors for predicting genetic structure are region and year.
While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
Aggressiveness and specific source field had little to no effect on genetic diversity. 
These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and severity [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables. 
This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.

>>>>

A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
We used this method to delineate the phenotypic (Aggressiveness, MCG), geographic (Region, Host, Location), and temporal (Year) components in influencing genetic composition of the populations.
<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data on the hierarchy of Region, Location, Host, Year as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr]. 
Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis. -->
To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

@@

<!-- 
 - Big Picture reminder of importance
   + This is an extension of a previous work [@otto-hanson2011variation]
   + White Mold Nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
   + It's unclear how genetic structure can influence
 - Critical Analysis of Major Findings
   + This is the largest genetic analysis of *S. sclerotiorum* to date
   + Similar to the findings of @otto-hanson2011variation, we found that
     Region has a strong influence on both population structure and
     aggressiveness.
   + North Dakota appeared
 - Additional Findings and how they fit into the literature
 - Limitations
 - Future Directions
 - Overall Conclusion and Impact
-->

@@

Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (Aggressiveness, Region, Source,   Year, and Host) [@legendre1999distance].
From the initial results, it appears that the most important factors for predicting genetic structure are region and year.
While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
Aggressiveness and specific source field had little to no effect on genetic diversity. 
These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables. 
This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.
